Legion: Expressing Locality and Independence with Logical Regions

### 2012-04-29 doc/sc2012/related.tex

Legion is an outgrowth of previous work on Sequoia, a locality-aware programming language capable
of expressing computations in deep memory hierarchies
\cite{Fatahalian06}.  Sequoia is a special case of the Legion
programming model in which only arrays can be recursively partitioned,
all access is exclusiv, there is a static mapping of tasks and data on to the machine (though
extensions to Sequoia make this mapping more dynamic \cite{BCSA11}) and, most fundamentally, the decomposition of tasks and the decomposition
of data is one-to-one.  Legion generalizes the Sequoia model by allowing for
partitioning of pointer data structures through regions, enabling dynamic mappings through
the mapper interface, and allowing different coherence
properties.  Legion's decoupling of the task tree from the region tree in Legion leads directly to the scheduling problem solved by our software out-of-order processor for tasks with region arguments.

@@

Deterministic Parallel Java (DPJ) is a parallel extension of Java
that, like Legion, uses regions to express locality and does
dependence analysis at the granularity of conflicts on region
arguments to functions \cite{Bocchino09}.  As the name suggests, one
of the primary goals of DPJ is to provide a programming model that
guarantees determinism while also providing supporting parallelism.
As a result the programming model is more restrictive than Legion, as
there is no notion of weaker coherence properties of regions than
exclusive access (which is essential to determinism).  DPJ can express
the populate-then-partition style of using regions, but not the
partition-then-populate; overall, the language is more static, and
thus provides more guarantees and less flexibility than Legion.  Most
significantly, however, DPJ efforts have so far focused on JVM
implementations on on shared memory machines; there is nothing
corresponding to our distributed software out-of-order processor or
the mapping interface.

@@

In previous programming systems regions have primarily been used as a construct for
describing memory management schemes \cite{REAPS02}\cite{RC01}  
or for enforcing safety policies \cite{CYCLONE01}.  In all cases, regions had
memory layout implications.  Logical regions in Legion enable the programmer to
describe locality and independence without any memory layout implications.

>>>>

Legion began as an outgrowth of Sequoia, a locality-aware programming language capable
of expressing computations in deep memory hierarchies
\cite{Fatahalian06}.  Sequoia is a special case of the Legion
programming model in which only arrays can be recursively partitioned,
all access is exclusive, there is a static mapping of tasks and data on to the machine (though
extensions to Sequoia make this mapping more dynamic \cite{BCSA11}) and, most fundamentally, the decomposition of tasks and the decomposition
of data is one-to-one.  Legion generalizes the Sequoia model by allowing for
partitioning of pointer data structures through regions, enabling dynamic mappings through
the mapper interface, and allowing different coherence
properties.  Legion's decoupling of the task tree from the region tree in Legion leads directly to the scheduling problem solved by our software out-of-order processor for tasks with region arguments.

@@

Deterministic Parallel Java (DPJ) is a parallel extension of Java
that, like Legion, uses regions to express locality and does
dependence analysis at the granularity of conflicts on region
arguments to functions \cite{Bocchino09}.  As the name suggests, one
of the primary goals of DPJ is to provide a programming model that
guarantees determinism while also supporting parallelism.
As a result the programming model is more restrictive than Legion, as
there is no notion of weaker coherence properties of regions than
exclusive access (which is essential to determinism).  DPJ can express
the populate-then-partition style of using regions, but not the
partition-then-populate; overall, the language is more static, and
thus provides more guarantees and less flexibility than Legion.  
DPJ efforts have so far focused on JVM implementations on shared memory machines.
There have also been several other recent efforts that propose task schedulers
based on dynamic detection of memory conflicts, but these also have assumed
underlying shared memory hardware \cite{VTN11,Tzenakis12,EYJD12}.

@@

In previous programming systems regions have primarily been used as a construct for
describing memory management schemes \cite{REAPS02,RC01}  
or for enforcing safety policies \cite{CYCLONE01}.  We follow \cite{RC01} in Legion's
decision to make regions first class.  In these works, however, regions have
memory layout implications.  Logical regions in Legion enable the programmer to
describe locality independent of memory layout.
